---
title: "SOP - Package `checkpoint`"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE,
                      eval = FALSE, message = FALSE)
library(emo)
library(knitr)
```


## Overview

Following the package's creator [description](https://github.com/RevolutionAnalytics/checkpoint):

> The goal of `checkpoint` is to solve the problem of package reproducibility in R. Specifically, `checkpoint` solve the problems that occur when you don’t have the correct versions of R packages. Since packages get updated on CRAN all the time, it can be difficult to recreate an environment where all your packages are consistent with some earlier state. To solve this, `checkpoint` allows you to install packages from a specific snapshot date. In other words, checkpoint makes it possible to install package versions from a specific date in the past, as if you had a CRAN time machine.

To get more information about the package functionality, please read the [package README](https://github.com/RevolutionAnalytics/checkpoint), available on github.

After you have read and understood what is happening behind the scenes, you can follow the step-by-step to run and enjoy the benefits of the package.


## Step-by-step


### 1. Create a R project

As you learned, `checkpoint` works best with R projects. To create a project in RStudio:

`File -> New Project...`


### 2. Open a R script

`File -> New file -> R script` or `Ctrl(Cmd) + Shift + N`

Save this script in your project directory.


### 3. Install and load `checkpoint`

#### From CRAN

```{r}
install.packages("checkpoint")
library(checkpoint)
```

#### From github repo (development version)

```{r}
install.packages("devtools")
devtools::install_github("RevolutionAnalytics/checkpoint")
library(checkpoint)
```


### 4. Write "checkpoint("YYYY-MM-DD")", in the top of the script

Remember: **DON'T RUN!**

```{r}
checkpoint("2021-06-10")
```


### 5. Add some packages

If you run the code, R will load the installed versions of the packages, but this is **not** what we want. So, don't run yet.

Remember: save your script.

```{r}
library(dplyr)
library(car)
library(readxl)
```


### 6. Run the checkpoint

When you run, `checkpoint` will look for `library()`, `require()`, `::` or `:::` commands in the scripts saved in your project directory.

If it finds something, the download and installation will start.


```{r}
checkpoint("2021-06-10") # Run
```


### 7. Check the .libPaths()

To see from where R is loading the packages now, run:

```{r}
.libPaths()
```

You will find a new folder `~/.checkpoint` in the top.


### 8. Now, you can work!

```{r}
install.packages("checkpoint")
library(checkpoint)

# checkpoint
checkpoint("2021-06-10")

# packages
library(dplyr)
library(car)
library(readxl)

# analysis
my_data <- readxl::read_excel("my-data.xlsx")
summary(my_data)
```



### 9. Oh! I forgot some packages `r emo::ji("person_facepalming")`

No worries! You can insert new packages, but remember:

1. write `library(<package>)` (don't run)

2. save the script

3. run `checkpoint`

4. load the package


```{r}
install.packages("checkpoint")
library(checkpoint)

# checkpoint
checkpoint("2021-06-10") # 3. run

# packages
library(dplyr)
library(car)
library(readxl)

# forgotten packages
library(ggplot2) # 1. write; 2. save; 4. run
library(broom)

# analysis
my_data <- readxl::read_excel("my-data.xlsx")
summary(my_data)
```


### 10. No more checkpoints today

If you want to stop `checkpoint`, you can:

- run `uncheckpoint()`

- restart R (`Ctrl(Cmd) + Shift + F10`)
